An elementary and constructive proof of Grothendieck's generic freeness lemma

### 2018-04-15 paper-generic-freeness.tex
Proof of the finitely-generated case

\section{The proof}

@@

>>>>

\section{The proof of the general case}

@@

\section{The proof of the finitely-generated case}

The following proposition is just an instance of Grothendieck's generic
freeness lemma. We prove it here because it admits an easier proof.

\begin{prop}Let~$A$ be a reduced ring. Let~$M$ be a finitely
generated~$A$-module. If~$f = 0$ is the only element of~$A$ such
that~$M[f^{-1}]$ is a finite free~$A[f^{-1}]$-module, then~$1 = 0$ in~$A$.
\end{prop}

\begin{proof}We proceed by induction on the length of a given generating family
of~$M$. Let~$M$ be generated by~$(v_1,\ldots,v_m)$.

We show that the family~$(v_1,\ldots,v_m)$ is linear independent. Let~$\sum_i
a_i v_i = 0$. Over~$A[a_i^{-1}]$, the vector~$v_i \in M[a_i^{-1}]$ is a linear
combination of the other generators. Thus~$M[a_i^{-1}]$ can be generated as
an~$A[a_i^{-1}]$-module by fewer than~$m$ generators. The induction hypothesis,
applied to this module, yields that~$1 = 0$ in~$A[a_i^{-1}]$. Since~$A$ is
reduced, this amounts to~$a_i = 0$.

We finish by using the assumption for~$f = 1$.
\end{proof}
